### 2024-02-10 paper/paper.qmd
Saved the data using the "01-download_data.R" file

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

[@citelabelled]

# Data {#sec-data}

# Results

# Discussion

\appendix

# Appendix {-}

# Additional data details

\newpage

# References