### 2024-02-11 other/ssrp_report/ssrp_report.qmd
Completed the SSRP Report

---
title: "SSRP Report based on *Who Do Voter ID Laws Keep from Voting?* by Fraga and Miller (2022)"
author: "Benny Rochwerg"
format: pdf
editor: visual
---

```{r}
#| echo: false
#| warning: false
#| message: false

#### Loading Packages ####

# install.packages("tidyverse")
# install.packages("patchwork")
# install.packages("here")
library(tidyverse)
library(patchwork)
library(here)
```

```{r}
#| echo: false
#| warning: false
#| message: false

# Loading the cleaned data
reasons_data_cleaned <- read_csv(here::here("data/replication/reasons_data_cleaned.csv"))
```

# Report

Using the R programming language (R Core Team 2023) and the tidyverse (Wickham et al. 2019), patchwork (Pedersen 2024), haven (Wickham, Miller, and Smith 2023), labelled (Larmarange 2023), and here (Müller 2020) packages, three figures from *Who Do Voter ID Laws Keep from Voting?* by Fraga and Miller (2022) were reproduced.

```{r}
#| echo: false
#| warning: false
#| label: fig-2
#| fig-cap: "Reproduction of Figure 2 from Fraga and Miller (2022)"

# Some of this code was adapted from Alexander (2023).

# Replication of Figure 2 from the paper (Fraga and Miller 2022)

# Creating a table containing the relevant voter ID reasons and percentages
figure_2_data <- tibble(
  reason = c("Family Obligations", "Applied, Has Not Received",
             "Lack of Transportation", "Lacks Necessary Documents",
             "Disability/Illness", "Work Obligations",
             "ID was Lost or Stolen", "Other Reason"),
  percent = round((100 * c(mean(reasons_data_cleaned$family),
                    mean(reasons_data_cleaned$applied),
                    mean(reasons_data_cleaned$transport),
                    mean(reasons_data_cleaned$birthcert),
                    mean(reasons_data_cleaned$disability),
                    mean(reasons_data_cleaned$work),
                    mean(reasons_data_cleaned$lost),
                    mean(reasons_data_cleaned$other))), 2))

# Creating data labels to be shown on the graph
figure_2_data_labels <- paste(
  figure_2_data$percent, "%", sep = "")

# Adding the data labels to the data table
figure_2_data <- figure_2_data |>
  mutate(label = figure_2_data_labels)

# Creating the replicated graph
figure_2_replication <-
  ggplot(mapping = aes(x = reorder(figure_2_data$reason, figure_2_data$percent,
                                   decreasing = TRUE),
                       y = figure_2_data$percent)) +
  geom_col() +
  theme_classic() +
  labs(x = "Impediment to voter ID", y = "Percentage",
       title = "Figure 2. Impediments cited by Texas voters in 2016") +
  coord_flip() +
  # The geom_text function used to add the labels to the graph was obtained from
  # R Graphics Cookbook:
  # https://r-graphics.org/recipe-bar-graph-labels
  # The hjust argument used to align the labels was obtained from
  # Cédric Scherer's website:
  # https://www.cedricscherer.com/2021/07/05/a-quick-how-to-on-labelling-bar-graphs-in-ggplot2/
  geom_text(aes(label = figure_2_data$label),
            hjust = c(0, 0, 0, 0, 0, 0, 0, 0),
            colour = "red",
            size = 3) +
  # The expand_limits function used to change the upper limit of the percent
  # axis's scale was obtained from the tidyverse ggplot2 website:
  # https://ggplot2.tidyverse.org/reference/expand_limits.html
  expand_limits(y = 40)

# Displaying the replicated graph
figure_2_replication
```

```{r}
#| echo: false
#| warning: false
#| label: fig-3
#| fig-cap: "Reproduction of Figure 3 from Fraga and Miller (2022)"

# Some of this code was adapted from Alexander (2023).

# Replication of Figure 3 from the paper (Fraga and Miller 2022)

# Creating a table containing the relevant voter ID reasons and percentages
figure_3_data <- tibble(
  reason = c("Relocation", "Hardship", "ID-Capable"),
  percent = round((100 * c(mean(reasons_data_cleaned$relocation),
                           mean(reasons_data_cleaned$hardship),
                           mean(reasons_data_cleaned$id_capable))), 2))

# Creating data labels to be shown on the graph
figure_3_data_labels <- paste(
  figure_3_data$percent, "%", sep = "")

# Adding the data labels to the data table
figure_3_data <- figure_3_data |>
  mutate(label = figure_3_data_labels)

# Creating the replicated graph
figure_3_replication <-
  ggplot(mapping = aes(x = reorder(figure_3_data$reason, figure_3_data$percent,
                                   decreasing = TRUE),
                       y = figure_3_data$percent)) +
  geom_col() +
  theme_classic() +
  labs(x = "Impediment to voter ID", y = "Percentage",
       title = "Figure 3. Binned impediments cited") +
  coord_flip() +
  # The geom_text function used to add the labels to the graph was obtained from
  # R Graphics Cookbook:
  # https://r-graphics.org/recipe-bar-graph-labels
  # The hjust argument used to align the labels was obtained from
  # Cédric Scherer's website:
  # https://www.cedricscherer.com/2021/07/05/a-quick-how-to-on-labelling-bar-graphs-in-ggplot2/
  geom_text(aes(label = figure_3_data$label),
            hjust = c(0, 0, 0),
            colour = "red",
            size = 3) +
  # The expand_limits function used to change the upper limit of the percent
  # axis's scale was obtained from the tidyverse ggplot2 website:
  # https://ggplot2.tidyverse.org/reference/expand_limits.html
  expand_limits(y = 40)

# Displaying the replicated graph
figure_3_replication
```

\newpage

```{r}
#| echo: false
#| warning: false
#| label: fig-a2
#| fig-cap: "Reproduction of Figure A2 from Fraga and Miller (2022)"

# Some of this code was adapted from Alexander (2023).

# Replication of Figure A2 from the paper (Fraga and Miller 2022)

# Filtering the cleaned dataset to only include entries with the race "White"
reasons_data_cleaned_white <- reasons_data_cleaned |>
  filter(race == "White")

# Creating a table containing the relevant voter ID reasons and percentages
# for the race "White"
figure_a2_data_white <- tibble(
  reason = c("Relocation", "Hardship", "ID-Capable"),
  percent = round((100 * c(mean(reasons_data_cleaned_white$relocation),
                           mean(reasons_data_cleaned_white$hardship),
                           mean(reasons_data_cleaned_white$id_capable))), 2))

# Creating data labels to be shown on the graph for the race "White"
figure_a2_data_white_labels <- paste(
  figure_a2_data_white$percent, "%", sep = "")

# Adding the data labels to the data table for the race "White"
figure_a2_data_white <- figure_a2_data_white |>
  mutate(label = figure_a2_data_white_labels)

# Creating the replicated graph for the race "White"
figure_a2_white_replication <-
  ggplot(mapping = aes(x = figure_a2_data_white$reason,
                       y = figure_a2_data_white$percent)) +
  geom_col() +
  theme_classic() +
  labs(x = "Impediment to voter ID", y = "Percentage") +
  scale_x_discrete(limits = c("ID-Capable", "Hardship", "Relocation")) +
  coord_flip() +
  # The geom_text function used to add the labels to the graph was obtained from
  # R Graphics Cookbook:
  # https://r-graphics.org/recipe-bar-graph-labels
  # The hjust argument used to align the labels was obtained from
  # Cédric Scherer's website:
  # https://www.cedricscherer.com/2021/07/05/a-quick-how-to-on-labelling-bar-graphs-in-ggplot2/
  geom_text(aes(label = figure_a2_data_white$label),
            hjust = c(0, 0, 0),
            colour = "red",
            size = 2) +
  ggtitle("White") +
  # The expand_limits function used to change the upper limit of the percent
  # axis's scale was obtained from the tidyverse ggplot2 website:
  # https://ggplot2.tidyverse.org/reference/expand_limits.html
  expand_limits(y = 45)

# Filtering the cleaned dataset to only include entries with the race "Black"
reasons_data_cleaned_black <- reasons_data_cleaned |>
  filter(race == "Black")

# Creating a table containing the relevant voter ID reasons and percentages
# for the race "Black"
figure_a2_data_black <- tibble(
  reason = c("Relocation", "Hardship", "ID-Capable"),
  percent = round((100 * c(mean(reasons_data_cleaned_black$relocation),
                           mean(reasons_data_cleaned_black$hardship),
                           mean(reasons_data_cleaned_black$id_capable))), 2))

# Creating data labels to be shown on the graph for the race "Black"
figure_a2_data_black_labels <- paste(
  figure_a2_data_black$percent, "%", sep = "")

# Adding the data labels to the data table for the race "Black"
figure_a2_data_black <- figure_a2_data_black |>
  mutate(label = figure_a2_data_black_labels)

# Creating the replicated graph for the race "Black"
figure_a2_black_replication <-
  ggplot(mapping = aes(x = figure_a2_data_black$reason,
                       y = figure_a2_data_black$percent)) +
  geom_col() +
  theme_classic() +
  labs(x = "Impediment to voter ID", y = "Percentage") +
  scale_x_discrete(limits = c("ID-Capable", "Hardship", "Relocation")) +
  coord_flip() +
  # The geom_text function used to add the labels to the graph was obtained from
  # R Graphics Cookbook:
  # https://r-graphics.org/recipe-bar-graph-labels
  # The hjust argument used to align the labels was obtained from
  # Cédric Scherer's website:
  # https://www.cedricscherer.com/2021/07/05/a-quick-how-to-on-labelling-bar-graphs-in-ggplot2/
  geom_text(aes(label = figure_a2_data_black$label),
            hjust = c(0, 0, 0),
            colour = "red",
            size = 2) +
  ggtitle("Black") +
  # The expand_limits function used to change the upper limit of the percent
  # axis's scale was obtained from the tidyverse ggplot2 website:
  # https://ggplot2.tidyverse.org/reference/expand_limits.html
  expand_limits(y = 45)

# Filtering the cleaned dataset to only include entries with the race "Latinx"
reasons_data_cleaned_latinx <- reasons_data_cleaned |>
  filter(race == "Latinx")

# Creating a table containing the relevant voter ID reasons and percentages
# for the race "Latinx"
figure_a2_data_latinx <- tibble(
  reason = c("Relocation", "Hardship", "ID-Capable"),
  percent = round((100 * c(mean(reasons_data_cleaned_latinx$relocation),
                           mean(reasons_data_cleaned_latinx$hardship),
                           mean(reasons_data_cleaned_latinx$id_capable))), 2))

# Creating data labels to be shown on the graph for the race "Latinx"
figure_a2_data_latinx_labels <- paste(
  figure_a2_data_latinx$percent, "%", sep = "")

# Adding the data labels to the data table for the race "Latinx"
figure_a2_data_latinx <- figure_a2_data_latinx |>
  mutate(label = figure_a2_data_latinx_labels)

# Creating the replicated graph for the race "Latinx"
figure_a2_latinx_replication <-
  ggplot(mapping = aes(x = figure_a2_data_latinx$reason,
                       y = figure_a2_data_latinx$percent)) +
  geom_col() +
  theme_classic() +
  labs(x = "Impediment to voter ID", y = "Percentage") +
  scale_x_discrete(limits = c("ID-Capable", "Hardship", "Relocation")) +
  coord_flip() +
  # The geom_text function used to add the labels to the graph was obtained from
  # R Graphics Cookbook:
  # https://r-graphics.org/recipe-bar-graph-labels
  # The hjust argument used to align the labels was obtained from
  # Cédric Scherer's website:
  # https://www.cedricscherer.com/2021/07/05/a-quick-how-to-on-labelling-bar-graphs-in-ggplot2/
  geom_text(aes(label = figure_a2_data_latinx$label),
            hjust = c(0, 0, 0),
            colour = "red",
            size = 2) +
  ggtitle("Latinx") +
  # The expand_limits function used to change the upper limit of the percent
  # axis's scale was obtained from the tidyverse ggplot2 website:
  # https://ggplot2.tidyverse.org/reference/expand_limits.html
  expand_limits(y = 45)

# Combining the graphs for the three races into one graph
figure_a2_replication <- (figure_a2_white_replication) / (figure_a2_black_replication) / (figure_a2_latinx_replication)

# Adding a title to the combined graph
figure_a2_replication <- figure_a2_replication +
  plot_annotation(title = "Figure A2. Impediments cited by Texas voters in 2016, by race")

# Combining the axis labels on the combined graph
figure_a2_replication <- figure_a2_replication +
  plot_layout(axes = "collect")

# Displaying the replicated graph
figure_a2_replication
```

@fig-2, @fig-3, and @fig-a2 suggest that the reproduction of these figures from Fraga and Miller (2022) was successful.

\newpage

# References

Alexander R. (2023). *Telling Stories with Data*. Boca Raton: CRC Press. <https://tellingstorieswithdata.com/>.

Fraga BL, Miller MG. (2022). "Who Do Voter ID Laws Keep from Voting?" *The Journal of Politics* 84 (2): 1091–1105. <https://doi.org/10.1086/716282>.

Larmarange J (2023). *labelled: Manipulating Labelled Data*. R package version 2.12.0, <https://CRAN.R-project.org/package=labelled>.

Müller K (2020). *here: A Simpler Way to Find Your Files*. R package version 1.0.1, <https://CRAN.R-project.org/package=here>.

Pedersen T (2024). *patchwork: The Composer of Plots*. R package version 1.2.0, <https://CRAN.R-project.org/package=patchwork>.

R Core Team (2023). *R: A Language and Environment for Statistical Computing*. R Foundation for Statistical Computing, Vienna, Austria. <https://www.R-project.org/>.

Wickham H, Averick M, Bryan J, Chang W, McGowan LD, François R, Grolemund G, Hayes A, Henry L, Hester J, Kuhn M, Pedersen TL, Miller E, Bache SM, Müller K, Ooms J, Robinson D, Seidel DP, Spinu V, Takahashi K, Vaughan D, Wilke C, Woo K, Yutani H (2019). “Welcome to the tidyverse.” *Journal of Open Source Software*, 4 (43), 1686. doi:10.21105/joss.01686 <https://doi.org/10.21105/joss.01686>.

Wickham H, Miller E, Smith D (2023). *haven: Import and Export 'SPSS', 'Stata' and 'SAS' Files*. R package version 2.5.4, <https://CRAN.R-project.org/package=haven>.